# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Loading the data file with a header. Ensure that dates are recognized as dates.
The csv file should be in the same folder than this Rmd file. 
You need to set this folder as the working directory. 

```{r, dataloading}
data = read.csv('activity.csv', h=T, na.strings = "NA")
data$date <- as.Date(data$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
Loading the 'plyr' package to compute the sum of steps per day.
Then, plot these sums as an histogram.

```{r, steps_per_day}
library(plyr)
# Keep only complete case
data.complete = data[ complete.cases(data), ]
sum_steps = ddply(data.complete, .(date), summarise, sumSteps=sum(steps))
hist(sum_steps$sumSteps, main='Histogram of total steps per day', xlab='Total of steps per day')
# Change the limit to display values as scientific format
options( scipen = 20 )
```

The mean total number of steps taken each day is **`r mean(sum_steps$sumSteps)`** and the median is **`r median(sum_steps$sumSteps)`**.


## What is the average daily activity pattern?
Compute the number of steps per interval of 5-minutes across all days.
Plot these averages using ggplot. 
Finally, find which interval corresponds to the maximum number of steps.

```{r, timeseries}
steps.interval = ddply(data.complete, .(interval), summarise, mSteps=mean(steps))
require(ggplot2)
ggplot( data = steps.interval, aes( interval, mSteps )) + geom_line() +
    xlab("5-minutes interval number") +
    ylab("Average number of steps")
maxinter = steps.interval[steps.interval$mSteps==max(steps.interval$mSteps),1]
```

The 5-minutes interval, on average across all the days in the dataset, containing the maximum number of steps is the **`r maxinter`th** interval.


## Imputing missing values

There are **`r nrow(data[ is.na(data[,1]), ])`** missing values in the data frame.

Replace missing values by the mean of the current 5-minutes interval across all days.
Plot the new data as an histogram.

```{r, missingval}
# Function to replace NA value by the mean of equivalent interval
replace.na = function(x){
  if( is.na(x[1]) ){
    value = steps.interval[ which(steps.interval$interval==as.numeric(x[3])), 2]
  }else{
    value=x[1]
  } 
  return(value)
}
# Apply the function to each row of the data frame
nsteps = as.numeric(apply(data, 1, replace.na))
# Combined the original data frame with the filled values
dtfilled = cbind(data, nsteps)
sum_dtfilled = ddply(dtfilled, .(date), summarise, sumSteps=sum(nsteps))
hist(sum_dtfilled$sumSteps, main='Histogram of total steps per day', xlab='Total of steps per day')
```

The mean total number of steps taken each day with NA replaced is **`r mean(sum_dtfilled$sumSteps)`** and the median is **`r median(sum_dtfilled$sumSteps)`**.
These values are very close to those compute by removing the missing-values. 
The means are equals and the the median with filled values is slighty below the one with  missing-values removed. 


## Are there differences in activity patterns between weekdays and weekends?

```{r, weekend}
# Subset only weekend days and add 'weekend' as label
weekend = subset(dtfilled, weekdays(date)=='Saturday' | weekdays(date)=='Sunday')
weekend$weekend = 'weekend'
# Subset only week days and add 'week' as label
weekday = subset(dtfilled, weekdays(date)!='Saturday' & weekdays(date)!='Sunday')
weekday$weekend = 'weekday'
# Recombined week and weekend data frame
data.week = rbind(weekend,weekday)
data.week.me = ddply(data.week, .(weekend, interval), summarise, mSteps=mean(nsteps))
# Create the time series plot by weekday kind
ggplot( data = data.week.me, aes( interval, mSteps )) + geom_line() + 
  facet_grid(weekend~.) +
    xlab("5-minutes interval number") +
    ylab("Average number of steps")
```

